---
title: "EDS241: Take Home Final"
author: "Alexandra Yousefivand"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
library(tidyverse)
library(here)
library(estimatr)
library(stargazer)
library(car)

options(scipen=999) # not scientific notation
```



```{r data, include=TRUE}
# Load data
data_raw <- read_csv(here("KM_EDS241.csv"))

data <- data_raw %>% 
  mutate(year = as.factor(year), 
         nearinc = as.factor(nearinc))
```


\noindent (a) Using the data for 1981, estimate a simple OLS regression of real house values on the indicator for being located near the incinerator in 1981. What is the house value “penalty” for houses located near the incinerator? Does this estimated coefficient correspond to the ‘causal’ effect of the incinerator (and the negative amenities that come with it) on housing values? Explain why or why not.


```{r subset 1981}
# subset data
data_1981 <- data %>% filter(year == 1981)
```

```{r lm_robust}
model <- lm_robust(formula = rprice ~ nearinc, data = data_1981)
summary(model)
```


```{r penalty}
penalty <- abs(round(model$coefficients[2]))
```



**The house value "penalty" for houses located near the incinerator is `r penalty`; in other words, on average, houses near the incinerator cost $`r penalty` less than houses not near the incinerator. The very low p-value indicates that this is a statistically significant result and this estimated coefficient correlates with price. This might correspond to the 'causal' effect of the incinerator; however, there are other variables that may contribute to the difference in housing prices as well, which implies the possibility of omitted variables bias.**
 




\noindent (b) Using the data for 1978, provide some evidence the location choice of the incinerator was not “random”, but rather selected on the basis of house values and characteristics. [Hint: in the 1978 sample, are house values and characteristics balanced by nearinc status?]


```{r subset 1978}
# subset data
data_1978 <- data %>% filter(year == 1978)
```

```{r 0/1}
data_nearinc <- data_1978 %>% filter(nearinc == 1)
data_not_nearinc <- data_1978 %>% filter(nearinc == 0)
```


```{r price mean diff}
# unadjusted mean difference 
nearinc_mean_price <- mean(data_nearinc$rprice)

not_nearinc_mean_price <- mean(data_not_nearinc$rprice)

difference_price <- not_nearinc_mean_price - nearinc_mean_price
difference_price
```

**Houses near the incinerator cost, on average, $`r abs(round(difference_price))` less than houses not near the incinerator.** 

```{r age mean diff}
# unadjusted mean difference 
nearinc_mean_age <- mean(data_nearinc$age)

not_nearinc_mean_age <- mean(data_not_nearinc$age)

difference_age <- not_nearinc_mean_age - nearinc_mean_age
difference_age
```

**Houses near the incinerator are, on average, `r abs(round(difference_age))` years older than houses not near the incinerator.** 


```{r rooms mean diff}
# unadjusted mean difference 
nearinc_mean_rooms <- mean(data_nearinc$rooms)

not_nearinc_mean_rooms <- mean(data_not_nearinc$rooms)

difference_rooms <- not_nearinc_mean_rooms - nearinc_mean_rooms
difference_rooms
```

**Houses near the incinerator have, on average, `r abs(round(difference_rooms, 2))` fewer rooms than houses not near the incinerator.**

```{r area mean diff}
# unadjusted mean difference 
nearinc_mean_area <- mean(data_nearinc$area)

not_nearinc_mean_area <- mean(data_not_nearinc$area)

difference_area <- not_nearinc_mean_area - nearinc_mean_area
difference_area
```

**Houses near the incinerator have, on average, `r abs(round(difference_area))` less square footage (of the house) than houses not near the incinerator.**

```{r land mean diff}
# unadjusted mean difference 
nearinc_mean_land <- mean(data_nearinc$land)

not_nearinc_mean_land <- mean(data_not_nearinc$land)

difference_land <- not_nearinc_mean_land - nearinc_mean_land
difference_land
```

**Houses near the incinerator have, on average, `r abs(round(difference_land))` less square footage (of the lot) than houses not near the incinerator.**



```{r mean diff using regression}
# unadjusted mean difference using linear regression
model_age   <- lm_robust(formula = age ~ nearinc, data = data_1978)
summary(model_age)

model_rooms <- lm_robust(rooms ~ nearinc, data = data_1978)
summary(model_rooms)

model_area  <- lm_robust(area ~ nearinc, data = data_1978)
summary(model_area)

model_land  <- lm_robust(land ~ nearinc, data = data_1978)
summary(model_land)
```

**Additionally, each of these coefficients (or mean difference values) are statistically significant (p = 0.05). The above evidence implies the location choice of the incinerator was not “random”, but rather selected on the basis of housing prices and characteristics.**





\noindent (c) Based on the observed differences in (b), explain why the estimate in (a) is likely to be biased downward (i.e., overstate the negative effect of the incinerator on housing values).

**The estimate in (a), which is based on the observed differences in (b), is likely to be biased downward because this value captures the impact of other characteristics related to housing price (such as the age and size of the home) other than location relative to the incinerator. Before construction of the incinerator in 1978, homes near the incinerator site were older, smaller, and cost less, on average. Because the previous estimate absorbs the affect of these housing characteristics, it is likely to overstate the negative effect of the incinerator on housing values.**


\noindent (d) Use a difference-in-differences (DD) estimator to estimate the causal effect of the incinerator on housing values without controlling for house and lot characteristics. Interpret the magnitude and sign of the estimated DD coefficient.

```{r diff diff}
diff_diff <- lm_robust(formula = rprice ~ nearinc, data = data)
summary(diff_diff)
```

**The DD estimator is `r round(diff_diff$coeff[[2]])`, which implies houses near the incinerator are worth, on average, $`r abs(round(diff_diff$coefficients[[2]]))` less than houses not near the incinerator.**



\noindent (e) Report the 95% confidence interval for the estimate of the causal effect on the incinerator in (d).

```{r}
conf_low <- diff_diff$conf.low[[2]]
conf_high <- diff_diff$conf.high[[2]]
```


There is a 95% probability that the estimate of the causal effect on the incinerator is between $`r round(diff_diff$conf.low[[2]])` and $`r round(conf_low <- diff_diff$conf.high[[2]])`.


\noindent (f) How does your answer in (d) changes when you control for house and lot characteristics? Test the hypothesis that the coefficients on the house and lot characteristics are all jointly equal to 0.

