---
title: "EDS241: Take Home Final"
author: "Alexandra Yousefivand"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
library(tidyverse)
library(here)
library(estimatr)
library(stargazer)
library(car)

```



```{r , include=TRUE}
# Load data
data <- read.csv(here("KM_EDS241.csv"))
```

7 variables:
  -`rprice`: sales price of house
  -`nearinc`: =1 if near incinerator
  -`age`: age of house
  -`land`: square footage of the lot
  -`area`: square footage of the house
  -`rooms`: number of rooms
  -`year`: 1978 or 1981

\noindent **(a) Using the data for 1981, estimate a simple OLS regression of real house values on the indicator for being located near the incinerator in 1981.**

```{r}
# subset data
data_1981 <- data %>% filter(year == 1981)
```

```{r}
model <- lm(data = data_1981, formula = rprice ~ nearinc)
summary(model)

model_robust <- lm_robust(data = data_1981, formula = rprice ~ nearinc)
summary(model_robust)
```


What is the house value “penalty” for houses located near the incinerator? Does this estimated coefficient correspond to the ‘causal’effect of the incinerator (and the negative amenities that come with it) on housing values? Explain why or why not.

The house value "penalty" for houses located near the incinerator is $`r summary(model)[2]`

```{r}

```


